Precision calculations for cosmic shear

### 2017-02-09 limber2.tex
MK discussion of full projection flat-sky Pgamma

@@

\input{authors.tex}

@@

\begin{document}
\setlength{\voffset}{-12mm}

@@

\maketitle
\begin{abstract}

@@

This paper is organised as follows.  In section~\ref{sec:wl} we provide a pedagogical introduction to weak gravitational lensing theory, projections and power spectra for the flat-sky and spherical cases.  In section~\ref{sec:L2} we derive weak lensing observables using a second-order limber approximation first introduced by \citet{2008PhRvD..78l3506L}.  We compare the shear power spectrum and commonly-used two-point shear correlation function for the full solution to a range of different approximations in section~\ref{sec:comp}, providing cosmological constraints for each case using CFHTLenS data from \citet{CFHTLenS-2pt-notomo} in section~\ref{sec:cfhtlens}.  This paper draws from several sources of literature which have previously reviewed the accuracy of the Limber approximation for weak lensing studies, namely \citet{2008PhRvD..78d3002S,2012PhRvD..86b3001B, 2012MNRAS.422.2854G, 2016arXiv161104954K}.   We present a discussion and comparison of our results to these papers in Appendix~\ref{app:B}.

@@

The lensing potential $\psi$ in the Born approximation is defined as the
projected 3D metric potential $\Phi$ along the line of sight
\citep{1998ApJ...498...26K,BS01},
%
% Hu00 (21,22) with minus sign
%
\begin{equation}
  \psi(\theta, \varphi) = \frac 2 {{\rm c}^2} \int_0^{\chi_{\rm lim}} \frac{{\rm d}\chi}{\chi}
    \Phi[\chi, f_K(\chi) \theta, f_K(\chi) \varphi; \chi] \, q(\chi),
  \label{eq:psi}
\end{equation}
%
where the lensing efficiency $q$ is given by
%
\begin{equation}
  q(\chi) = \int\limits_\chi^{\chi_{\rm lim}} {\rm d} \chi^\prime \, n(\chi^\prime)
    \frac{f_K(\chi^\prime - \chi)}{f_K(\chi^\prime)},
  %
  \label{eq:lens_efficiency}
\end{equation}
%
corresponding to a population of lensed galaxies with normalised source redshift
distribution $n(z) {\rm d}z = n(\chi) {\rm d} \chi$. Here, ${\rm c}$ is the
speed of light, the projection is carried out over comoving distances $\chi$ up
to a limiting distance $\chi_{\rm lim}$, and $f_K$ is the the comoving angular
distance, which depend on the spatial curvature of the Universe $K$.
%
This assumes a homogeneous galaxy distribution without clustering, so that the
redshift distribution in this approximation does not depend on the position
$(\theta, \varphi)$. Accounting for this position dependence leads to
correction of weak-lensing quantities due to clustering of source galaxies with
other sources \citep{2002A&A...389..729S}, and with galaxies associated with
lens structures \citep{1998A&A...338..375B,H02}.

The 3D potential is related to the density contrast $\delta$ via the Poisson
equation. Assuming General Relativity, this relation is written in Fourier space as
%
\begin{align}
  \hat \Phi(\vec k; \chi) = & - \frac 3 2 \Omega_{\rm m} H_0^2 k^{-2} a^{-1}(\chi) \hat \delta(\vec k; \chi),
      \label{eq:poisson}
\end{align}
%
where $\Omega_{\rm m}$ is the matter density parameter, $H_0$ the Hubble constant, $\vec k$ a 3D Fourier wave
vector, and $a$ the scale factor with $a=1$ today.
The Fourier transform of the potential and its inverse are defined as
%
\begin{align}
  \hat \Phi(\vec k; \chi) = &  \int {\rm d}^3 r \, \Phi(\vec r; \chi) {\rm e}^{{\rm i} \vec k \vec r} .
  \label{eq:hatPhi}
  \\
  \Phi(\vec r; \chi) = &  \int \frac{{\rm d}^3 k}{(2\pi)^3}
      \hat \Phi(\vec k; \chi) {\rm e}^{-{\rm i} \vec r \vec k},
  \label{eq:hatPhi_inv}
\end{align}
%
where the integration range for both integrals is $\mathbb{R}^3$.

@@

Following \cite{2000PhRvD..62d3007H} we decompose the lensing potential
(\ref{eq:psi}) into a spherical harmonics expansion, in analogy of CMB
temperature, both of which are scalar functions on the sphere. This
decomposition and its inverse are
%
% Hu00 (23)
%
\begin{align}
  \psi(\theta, \varphi) = \sum_{\ell=0}^\infty \sum_{m=-\ell}^\ell \psi_{\ell m} {\rm Y}_{\ell m}(\theta, \varphi);
    \label{eq:psi_harm_exp}
    \\
    %
  \psi_{\ell m} = \int_{\mathbb{S}^2} {\rm d} \Omega \, \psi(\theta, \varphi) {\rm Y}^\ast_{\ell m}(\theta, \varphi).
  \label{eq:psi_harm_exp_inv}
\end{align}

To specify tomographic redshift bins $i=0\ldots N_z-1$, we introduce a family
of lensing efficiency functions $q_i$ defined by a corresponding family of
redshift distributions $n_i$ via eq.~(\ref{eq:lens_efficiency}). The resulting
lensing potential is denoted by $\psi_{\ell m, i}$. The tomographic power
spectrum of the lensing potential between two redshift bins $i$ and $j$,
$C_{ij}^\psi$ is then defined by \cite{pee80}
%
\begin{equation}
  \left\langle \psi^{}_{\ell m, i} \, \psi^\ast_{\ell^\prime m^\prime, j} \right\rangle
    = \delta_{\ell \ell^\prime} \delta_{m m^\prime} C^\psi_{ij}(\ell) .
  \label{eq:C_ell_psi}
\end{equation}
%
Using the properties of the spherical harmonics (see
App.~\ref{sec:derivations_C} for details) the power spectrum can be written as
%
\begin{align}
  C_{ij}^\psi(\ell) = & \frac 8 {c^4 \pi} 
  \int_0^{\chi_{\rm lim}} \frac{{\rm d}\chi}{\chi} q_i(\chi)
  \int_0^{\chi^\prime_{\rm lim}} \frac{{\rm d}\chi^\prime}{\chi^\prime} q_j(\chi^\prime)
  \int {\rm d} k k^2 \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_\Phi(k; \chi, \chi^\prime)
  \label{eq:C_ell_phi_Pphi} \\
  = & \frac 8 \pi \pref^2
  \int_0^{\chi_{\rm lim}} \frac{{\rm d}\chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
  \int_0^{\chi^\prime_{\rm lim}} \frac{{\rm d}\chi^\prime}{\chi^\prime} \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
  \int \frac{{\rm d} k}{k^2} \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_{\rm m}(k; \chi, \chi^\prime) ;
  \label{eq:C_ell_phi_Pm}
\end{align}
%
where ${\rm j}_\ell$ it the spherical Bessel function of order $\ell$. For
convenience we defined the normalisation constant $\pref$ as
%
\begin{equation}
  \pref = \frac 3 2 \Omega_{\rm m} \left(\frac{H_0} c\right)^2.
  \label{eq:pref}
\end{equation}
%
The first line expresses $C_{ij}^\psi$ in terms of the 3D potential power spectrum $P_\Phi$, defined as
%
\begin{align}
  \left\langle \hat \Phi(\vec k; \chi) \hat \Phi^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_\Phi(k; \chi, \chi^\prime).
    %\nonumber \\
    %= & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) \pref^2 k^{-4} a^{-1}(\chi) a^{-1}(\chi^\prime)
      %P_{\rm m}(k; \chi, \chi^\prime).
  \label{eq:p_phi}
\end{align}
%
The second line uses the 3D matter density power spectrum $P_{\rm m}$, which is defined analogously as
%
\begin{align}
  \left\langle \hat \delta(\vec k; \chi) \hat \delta^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_{\rm m}(k; \chi, \chi^\prime),
  \label{eq:p_m}
\end{align}
%
and is
related to $P_\Phi$ via the absolute square of the Poisson equation (\ref{eq:poisson}).

@@

Most previous work has used lensing quantities approximated on a flat sky,
neglecting the sky curvature. This is a valid approach for past and current
survey areas that have extents up to 10 degrees or less. The correlation
functions from the observed shear have been calculated using spherical
coordinates, and the effect on large scales is not negligible \cite{FSHK08}.
Here we examine the effect of sky curvature on the theoretical models.

@@

The power spectrum, the flat-sky version of
(\ref{eq:C_ell_psi}) is defined by
%
\begin{equation}
  \left\langle \hat \psi_i^{}(\vec \ell) \hat \psi_j^\ast(\vec \ell^\prime) \right\rangle
    = (2\pi)^2 \delta_{\rm D}(\vec \ell - \vec \ell^\prime) P^\psi_{ij}(\ell).
  \label{eq:P_ell_psi}
\end{equation}

\mk{I can derive a flat-sky shear power spectrum by replacing $\ellbar^2(\ell, 2)$ with $\ell^4$
in (\ref{eq:C_ell_full}), equivalent to replacing the edth operators with Euclidian derivatives.
However, this results in a different expression compared to other references, e.g.~\cite{2008PhRvD..78d3002S},
which is the following equation.}

The flat-sky power spectrum is
%
\begin{align}
  P_{ij}^\gamma(\ell) = & \frac 2 \pi \, \pref^2
                 \int_{0}^{\chi_{\rm lim}} {\rm d} \chi \chi \, \frac{q_i(\chi)}{a(\chi)}
                \int_{0}^{\chi_{\rm lim}} {\rm d} \chi^\prime\, \chi^\prime
                \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
                \nonumber \\
                & \times \int_0^\infty {\rm d} k \, k^2 \, P_{\rm m}(k; \chi, \chi^\prime) \,
                {\rm j}_\ell(k \chi) \, {\rm j}_\ell(k \chi^\prime) .
  \label{eq:P_ell_gamma}
\end{align}
%

\mk{I can get this by placing the prefactor under the integral and
setting $\ell^4 = k^4 \chi^2 {\chi^\prime}^2$. This does not seem to be the proper way.
Other derivations of the shear power spectrum (e.g.\cite{BS01} also assume other approximations
such as small-angle and Limber, but I do not know how to derive this without Limber.}

@@

Most cosmic shear analysis to date have been performed on real-space
correlation statistics, since these can be measured direcaly from an observed
galaxy shape catalogue. The basic quantity is the two-point correlation
function \citep{miraldaescude:1991}, which is given by
%
\begin{eqnarray}
  \xi_+(\theta) 
  % &
  = \frac 1 {2\pi} \int {\rm d} \ell \, \ell {\rm J}_0(\ell
   \theta)
  P^\gamma(\ell);
  \quad
  %\nonumber \\
   %
   \xi_-(\theta)
  %&
  = \frac 1 {2\pi} \int
   {\rm d} \ell \, \ell {\rm J}_4(\ell \theta)
  P^\gamma(\ell).
  %
   \label{eqn:xiGG}
\end{eqnarray}
%
where ${\rm J}_\nu$ first-kind Bessel function of order $\nu$. The flat-sky
shear power spectrum $P^\gamma$ an be related to the underlying matter power
spectrum through equation \ref{eq:P_ell_limber1} when adopting a first-order
extended Limber approximation, or equation~\ref{eq:P_ell_limber2} when adopting
a second-order extended Limber approximation \citep[for more details see][and
references therein]{miraldaescude:1991, kaiser:1992,
bartelmann/schneider:2001}.

@@

  \caption{Shear correlation functions $\xi_+$ (\emph{left}) and $\xi_-$ (\emph{right}).
    The lower panels shows the relative differences to the case ExtL2Sph,  see Table \ref{tab:cases} for the different
    approximations and their symbols.
    The theoretical model corresponds to the CFHTLenS best-fit parameters with
    $\Omega_{\rm m} = 0.279$, $h=0.701$, and $\sigma_8 = 0.79$ \citep{CFHTLenS-2pt-notomo},
    except the red dashed curve with adapts a Planck-best fit
    cosmology with $\Omega_{\rm m} = 0.3$, $h=0.67$ and $\sigma_8 = 0.83$ \citep{2015arXiv150201589P}.
  }

@@

The authors thank Tommaso Giannantonio, Sarah Bridle, Ami Choi, Donnacha Kirk, Lance Miller, Benjamin Joachimi, Chris Blake and Adam Amara for very helpful discussions.
This work was financially supported by the DFG (Emmy Noether grants Hi 1495/2-1 and SI 1769/1-1; TR33 `The Dark Universe'), the Alexander von Humboldt Foundation, the ERC (grants 279396, 647112), the Seventh Framework Programme of the European Commission (Marie Sklodwoska Curie Fellowship grant 656869), NSERC and CIfAR.

\ch{Check your acknowledgements}

@@

\subsection{Schmidt (2009)}

In eq.~(9) \cite{2008PhRvD..78d3002S} write the lensing power spectrum in the
flat-sky limit. Inserting the Poisson factor $D_\phi$, 
 the growth function $D_{\rm m}$, and writing the redshift filter
function $W_\kappa$ (eq.~10) in terms of the lensing efficiency and comoving distances,
$W_\kappa[z(\chi)] = H(z)^{-1} \chi q(\chi)$, we
%
%\begin{align}
  %W_\kappa(z) = & \frac 1 {H(z)} \int_z^\infty {\rm d} z_{\rm s} W_{\rm L}(z_{\rm s}, z) W_{\rm g}(z)
              %=  \frac 1 {H(z)} \chi \int_\chi^\infty {\rm d} \chi^\prime \frac{\chi^\prime - \chi}{\chi^\prime}
                %n(\chi)
              %\nonumber \\
              %& \equiv \frac 1 {H(z)} \chi q(\chi),
%\end{align}
%%
%using ${\rm d} z W_{\rm g}(z) = {\rm d} \chi n(\chi)$.
%To write the redshift integrals (9) in \cite{2008PhRvD..78d3002S} over 
%comoving distance, we make the variable transformation
%%
%\begin{equation}
  %{\rm d} \chi = \frac{c {\rm d} z}{H(z)} .
%\end{equation}
%%
%We
recover our (\ref{eq:P_ell_gamma}).

\subsection{Giannantonio et al. (2012)}

In eqs.~(25) and (26) \cite{2012MNRAS.422.2854G} write the flat-sky lensing
power spectrum. Their window function $W^{\varepsilon_i}$ defined in (25) for a
flat Universe equals $\pref \times q(\chi) / a(\chi)$, since ${\rm d} z \,
({\rm d} N(z)/{\rm d} z) = {\rm d} \chi n(\chi)$; however due to a typo there
is a factor $r_K[r(z)]$ missing in the window function, which translates into a
missing $r_K[r] r_K[r^\prime]$ in the full projection integral\footnote{Typo confirmed by T.~G.,
priv.~comm.}. This also leads to an errorous $r_K^{-2}(r)$ in the Limber
equation (27). With these factors accounted for, we reproduce the expressions
of \cite{2012MNRAS.422.2854G}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\author[M.~Kilbinger et~al.]
 {
  \parbox[h]{\textwidth}
  {
      Martin Kilbinger$^{1,2}$\thanks{E-mail: martin.kilbinger@cea.fr},
      %Catherine Heymans,
      %Peter Schneider,
      %Shahab Joudaki,
      %Marika Asgari,   % note that we'll worry about order at the end - Marika and Shahab put a lot of work into the comment paper, so we might just want to go alpha
      and other members of CFHTLenS and KiDS
      $\ldots$,
  }
  % 
  \vspace*{10pt} \\
  \hspace{-.1cm}$^1$ CEA/Irfu/SAp Saclay, Laboratoire AIM, 91191 Gif-sur-Yvette, France\\
  \hspace{-.1cm}$^2$ Institut d'Astrophysique de Paris, UMR7095 CNRS,
           Universit\'e Pierre \& Marie Curie, 98 bis boulevard Arago, 75014 Paris,
           France \\
 }

\voffset-0.50in

@@

\maketitle

@@

\begin{abstract}

@@

This paper is organised as follows.  In section~\ref{sec:wl} we provide a
pedagogical introduction to weak gravitational lensing theory, projections and
power spectra for the flat-sky and spherical cases, following the seminal work
from \citet{2000PhRvD..62d3007H}, see also \citet{2005PhRvD..72b3516C}. In section~\ref{sec:L2} we
derive weak lensing observables using a second-order limber approximation first
introduced by \citet{2008PhRvD..78l3506L}. We compare the shear power spectrum
and commonly-used two-point shear correlation function for the full solution to
a range of different approximations in section~\ref{sec:comp}, providing
cosmological constraints for each case using CFHTLenS data from
\citet{CFHTLenS-2pt-notomo} in section~\ref{sec:cfhtlens}. This paper draws
from several sources of literature which have previously discussed the full
projection, or reviewed the accuracy of the Limber approximation for weak
lensing studies, namely \citet{2008PhRvD..78d3002S,2012PhRvD..86b3001B,
2012MNRAS.422.2854G, 2016arXiv161104954K}. We present a discussion and
comparison of our results to these papers in Appendix~\ref{app:B}.

@@

The lensing potential $\psi$ at a direction on the sky $(\theta, \varphi)$ in
the Born approximation is defined as the projected 3D metric potential $\Phi$
along the line of sight \citep{1998ApJ...498...26K,BS01},
%
% Hu00 (21,22) with minus sign
%
\begin{equation}
  \psi(\theta, \varphi) = \frac 2 {{\rm c}^2} \int_0^{\chi_{\rm lim}} \frac{{\rm d}\chi}{\chi}
    \Phi[\chi, f_K(\chi) \theta, f_K(\chi) \varphi; \chi] \, q(\chi),
  \label{eq:psi}
\end{equation}
%
where the lensing efficiency $q$ is given by
%
\begin{equation}
  q(\chi) = \int\limits_\chi^{\chi_{\rm lim}} {\rm d} \chi^\prime \, n(\chi^\prime)
    \frac{f_K(\chi^\prime - \chi)}{f_K(\chi^\prime)},
  %
  \label{eq:lens_efficiency}
\end{equation}
%
corresponding to a population of lensed galaxies with normalised source
redshift distribution $n(z) {\rm d}z = n(\chi) {\rm d} \chi$. Here, ${\rm c}$
is the speed of light, the projection is carried out over comoving distances
$\chi$ up to a limiting distance $\chi_{\rm lim}$, and $f_K(\chi) = |K|^{-1/2}
S_K(|K|^{1/2} \chi)$ is the the comoving angular distance, where the function
$S_K$ equals $\sin$, ${\rm id}$, or $\sinh$ if the spatial curvature of the Universe
$K$ is positive, zero, or negative, respectively.

The form of the lensing efficiency $q$ (\ref{eq:lens_efficiency}) assumes a
homogeneous galaxy distribution without clustering, so that the redshift
distribution in this approximation does not depend on the direction on the sky.
Accounting for this position-dependence leads to correction of weak-lensing
quantities due to clustering of source galaxies with other sources
\citep{2002A&A...389..729S}, and with galaxies associated with lens structures
\citep{1998A&A...338..375B,H02}.

The 3D potential is related to the density contrast $\delta$ via the Poisson
equation. Assuming General Relativity, this relation is written in Fourier space as
%
\begin{align}
  \hat \Phi(\vec k; \chi) = & - \frac 3 2 \Omega_{\rm m} H_0^2 k^{-2} a^{-1}(\chi) \hat \delta(\vec k; \chi),
      \label{eq:poisson}
\end{align}
%
where $\Omega_{\rm m}$ is the matter density parameter, $H_0$ the Hubble constant, $\vec k$ a 3D Fourier wave
vector, and $a$ the scale factor with $a=1$ today.
The Fourier transform of the potential and its inverse are defined as
%
\begin{align}
  \hat \Phi(\vec k; \chi) = &  \int {\rm d}^3 r \, \Phi(\vec r; \chi) {\rm e}^{{\rm i} \vec k \vec r} ;
  \label{eq:hatPhi}
  \\
  \Phi(\vec r; \chi) = &  \int \frac{{\rm d}^3 k}{(2\pi)^3}
      \hat \Phi(\vec k; \chi) {\rm e}^{-{\rm i} \vec r \vec k},
  \label{eq:hatPhi_inv}
\end{align}
%
where the integration range for both integrals is $\mathbb{R}^3$.

@@

Following \cite{2000PhRvD..62d3007H} we decompose the lensing potential
(\ref{eq:psi}) into a spherical harmonics expansion, in analogy of CMB
temperature, both of which are scalar functions on the sphere. This
decomposition and its inverse are
%
% Hu00 (23)
%
\begin{align}
  \psi(\theta, \varphi) = \sum_{\ell=0}^\infty \sum_{m=-\ell}^\ell \psi_{\ell m} {\rm Y}_{\ell m}(\theta, \varphi);
    \label{eq:psi_harm_exp}
    \\
    %
  \psi_{\ell m} = \int_{\mathbb{S}^2} {\rm d} \Omega \, \psi(\theta, \varphi) {\rm Y}^\ast_{\ell m}(\theta, \varphi).
  \label{eq:psi_harm_exp_inv}
\end{align}
%
To specify tomographic redshift bins $i=0\ldots N_z-1$, we introduce a family
of lensing efficiency functions $q_i$ defined by a corresponding family of
redshift distributions $n_i$ via eq.~(\ref{eq:lens_efficiency}). The resulting
lensing potential is denoted by $\psi_{\ell m, i}$. The tomographic power
spectrum of the lensing potential between two redshift bins $i$ and $j$,
$C_{ij}^\psi$ is then defined by \cite{pee80}
%
\begin{equation}
  \left\langle \psi^{}_{\ell m, i} \, \psi^\ast_{\ell^\prime m^\prime, j} \right\rangle
    = \delta_{\ell \ell^\prime} \delta_{m m^\prime} C^\psi_{ij}(\ell) .
  \label{eq:C_ell_psi}
\end{equation}
%
Using the properties of the spherical harmonics (see
App.~\ref{sec:derivations_C} for details) the power spectrum can be written as
%
\begin{align}
  C_{ij}^\psi(\ell) = & \frac 8 {c^4 \pi} 
  \int_0^{\chi_{\rm lim}} \frac{{\rm d}\chi}{\chi} q_i(\chi)
  \int_0^{\chi^\prime_{\rm lim}} \frac{{\rm d}\chi^\prime}{\chi^\prime} q_j(\chi^\prime)
  \int {\rm d} k k^2 \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_\Phi(k; \chi, \chi^\prime)
  \label{eq:C_ell_phi_Pphi} \\
  = & \frac 8 \pi \pref^2
  \int_0^{\chi_{\rm lim}} \frac{{\rm d}\chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
  \int_0^{\chi^\prime_{\rm lim}} \frac{{\rm d}\chi^\prime}{\chi^\prime} \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
  \int \frac{{\rm d} k}{k^2} \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_{\rm m}(k; \chi, \chi^\prime) ;
  \label{eq:C_ell_phi_Pm}
\end{align}
%
where ${\rm j}_\ell$ it the spherical Bessel function of order $\ell$. For
convenience we defined the normalisation constant $\pref$ as
%
\begin{equation}
  \pref = \frac 3 2 \Omega_{\rm m} \left(\frac{H_0} c\right)^2.
  \label{eq:pref}
\end{equation}
%
The first line expresses $C_{ij}^\psi$ in terms of the 3D potential power spectrum $P_\Phi$, defined as
%
\begin{align}
  \left\langle \hat \Phi(\vec k; \chi) \hat \Phi^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_\Phi(k; \chi, \chi^\prime).
    %\nonumber \\
    %= & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) \pref^2 k^{-4} a^{-1}(\chi) a^{-1}(\chi^\prime)
      %P_{\rm m}(k; \chi, \chi^\prime).
  \label{eq:p_phi}
\end{align}
%
The second line uses the 3D matter density power spectrum $P_{\rm m}$, which is defined analogously as
%
\begin{align}
  \left\langle \hat \delta(\vec k; \chi) \hat \delta^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_{\rm m}(k; \chi, \chi^\prime),
  \label{eq:p_m}
\end{align}
%
and is
related to $P_\Phi$ via the absolute square of the Poisson equation (\ref{eq:poisson}).

@@

Most previous work has used lensing quantities approximated on a flat sky,
neglecting the sky curvature. This is a valid approach for past and current
survey areas that have extents up to 10 degrees or less. The correlation
functions from the observed shear have been calculated using spherical
coordinates, and the effect on large scales is not negligible \citep{FSHK08}.
Here we examine the effect of sky curvature on the theoretical models.

@@

The flat-sky power spectrum, the flat-sky analogon of (\ref{eq:C_ell_psi}) is defined by
%
\begin{equation}
  \left\langle \hat \psi_i^{}(\vec \ell) \hat \psi_j^\ast(\vec \ell^\prime) \right\rangle
    = (2\pi)^2 \delta_{\rm D}(\vec \ell - \vec \ell^\prime) P^\psi_{ij}(\ell).
  \label{eq:P_ell_psi}
\end{equation}
%
\cite{2000PhRvD..62d3007H} have shown that for small angles the harmonics
expansion (\ref{eq:psi_harm_exp}) can be approximated by the the Fourier
representation (\ref{eq:psi_fourier}). They also demonstrated that the power
spectra are approximately equal, $P^\psi \approx C^\psi$.

For a spin-2 field, \cite{2000PhRvD..62d3007H} approximates the edth operator by Cartesian derivatives, and approximates
(\ref{eq:sYlm_def}) as
%
\begin{equation}
  \ell^2 \, _{\pm2}\!{\rm Y}_{\ell m}(\theta, \varphi) =  {\rm e}^{{\mp}2 {\rm i} \phi_\ell}
    (\partial_1 \pm \partial_2)^2 \, {\rm Y}_{\ell m}(\theta, \varphi).
  \label{eq:sYlm_def_flat}
\end{equation}
%
The spherical prefactor $\sqrt{(\ell-1) \ell (\ell+1) (\ell+2)}$ is replaced by $\ell^2$ in flat coordiates.
We find for the flat-sky shear power spectrum
%
\begin{align}
  P^\gamma_{ij}(\ell) 
                 = & \frac 2 \pi \, \ell^4 \, \pref^2
                 \int_{0}^{\chi_{\rm lim}} \frac{\rm d \chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
                \int_{0}^{\chi_{\rm lim}} \frac{\rm d \chi^\prime}{\chi^\prime}
                \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
                %\nonumber \\
                %& \times
                \int_0^\infty \frac{{\rm d} k}{k^2} \, P_{\rm m}(k, \chi, \chi^\prime) \,
                {\rm j}_\ell(k \chi) \, {\rm j}_\ell(k \chi^\prime) .
  \label{eq:P_ell_gamma_full_der}
\end{align}
%
with flat-sky prefactor $p(\ell) = \ell^4$.
\mk{See App.~\ref{sec:schmidt08} and \ref{sec:giannantonio12} for discussions of alternative expressions of the flat-sky power spectrum.}

@@

Most cosmic shear analysis to date have been performed on real-space
correlation statistics, since these can be measured direcaly from an observed
galaxy shape catalogue. The basic quantity is the two-point correlation
function \citep{1991ApJ...370....1M}, which is given by
%
\begin{align}
  \xi_+(\theta) 
  % &
  = \frac 1 {2\pi} \int {\rm d} \ell \, \ell {\rm J}_0(\ell
   \theta)
  P^\gamma(\ell);
  \quad
  %\nonumber \\
   %
   \xi_-(\theta)
  %&
  = \frac 1 {2\pi} \int
   {\rm d} \ell \, \ell {\rm J}_4(\ell \theta)
  P^\gamma(\ell).
  %
   \label{eqn:xiGG}
\end{align}
%
where ${\rm J}_\nu$ first-kind Bessel function of order $\nu$. The flat-sky
shear power spectrum $P^\gamma$ an be related to the underlying matter power
spectrum through equation \ref{eq:P_ell_limber1} when adopting a first-order
extended Limber approximation, or equation~\ref{eq:P_ell_limber2} when adopting
a second-order extended Limber approximation \citep[for more details see][and
references therein]{1991ApJ...370....1M,1992ApJ...388..272K,
BS01}.

@@

  \caption{Shear correlation functions $\xi_+$ (\emph{left}) and $\xi_-$ (\emph{right}).
    The lower panels shows the relative differences to the case ExtL2Sph,  see Table \ref{tab:cases} for the different
    approximations and their symbols.
    The theoretical model corresponds to the CFHTLenS best-fit parameters with
    $\Omega_{\rm m} = 0.279$, $h=0.701$, and $\sigma_8 = 0.79$ \citep{2015arXiv150201589P},
    except the red dashed curve with adapts a Planck-best fit
    cosmology with $\Omega_{\rm m} = 0.3$, $h=0.67$ and $\sigma_8 = 0.83$ \citep{2015arXiv150201589P}.
  }

@@

The authors thank Tommaso Giannantonio, Sarah Bridle, Ami Choi, Donnacha Kirk, Lance Miller, Benjamin Joachimi, Chris Blake and Adam Amara for very helpful discussions.

@@

\subsection{Schmidt (2009)}
\label{sec:schmidt08}

In eq.~(9) \cite{2008PhRvD..78d3002S} write the lensing power spectrum in the
flat-sky limit. Inserting the Poisson factor $D_\phi$, 
 the growth function $D_{\rm m}$, and writing the redshift filter
function $W_\kappa$ (eq.~10) in terms of the lensing efficiency and comoving distances,
$W_\kappa[z(\chi)] = H(z)^{-1} \chi q(\chi)$, their expression, using our notations, reads
%
\begin{align}
  P_{ij}^\gamma(\ell) = & \frac 2 \pi \, \pref^2
                 \int_{0}^{\chi_{\rm lim}} {\rm d} \chi \chi \, \frac{q_i(\chi)}{a(\chi)}
                \int_{0}^{\chi_{\rm lim}} {\rm d} \chi^\prime\, \chi^\prime
                \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
                \nonumber \\
                & \times \int_0^\infty {\rm d} k \, k^2 \, P_{\rm m}(k; \chi, \chi^\prime) \,
                {\rm j}_\ell(k \chi) \, {\rm j}_\ell(k \chi^\prime) ,
  \label{eq:P_ell_gamma}
\end{align}
%
This is consistent with our equation (\ref{eq:P_ell_gamma_full_der}) under the additional assumption that
mainly modes with $k \chi \approx k \chi^\prime \approx \ell$ contribute to the integral; that is modes
around the maxima of the Bessel functions. Then, we can draw out of the integral the factor $\ell^4 \approx k^4 \chi^2 {\chi^\prime}^2$,
to recover (\ref{eq:P_ell_gamma_full_der}).
\mk{This argument is very hand-wavey! In fact, using the approximation $k \chi = \ell$ seems to go too far, this is already halfway the Limber approximation.}

%\begin{align}
  %W_\kappa(z) = & \frac 1 {H(z)} \int_z^\infty {\rm d} z_{\rm s} W_{\rm L}(z_{\rm s}, z) W_{\rm g}(z)
              %=  \frac 1 {H(z)} \chi \int_\chi^\infty {\rm d} \chi^\prime \frac{\chi^\prime - \chi}{\chi^\prime}
                %n(\chi)
              %\nonumber \\
              %& \equiv \frac 1 {H(z)} \chi q(\chi),
%\end{align}
%%
%using ${\rm d} z W_{\rm g}(z) = {\rm d} \chi n(\chi)$.
%To write the redshift integrals (9) in \cite{2008PhRvD..78d3002S} over 
%comoving distance, we make the variable transformation
%%
%\begin{equation}
  %{\rm d} \chi = \frac{c {\rm d} z}{H(z)} .
%\end{equation}
%%
%We

\subsection{Giannantonio et al. (2012)}
\label{sec:giannantonio12}

In eqs.~(25) and (26) \cite{2012MNRAS.422.2854G} write the flat-sky lensing
power spectrum. Their window function $W^{\varepsilon_i}$ defined in (25) for a
flat Universe equals $\pref \times q(\chi) / a(\chi)$, since ${\rm d} z \,
({\rm d} N(z)/{\rm d} z) = {\rm d} \chi n(\chi)$; however due to a typo there
is a factor $r_K[r(z)]$ missing in the window function, which translates into a
missing $r_K[r] r_K[r^\prime]$ in the full projection integral\footnote{Typo
confirmed by T.~G., priv.~comm.}. This also leads to an errorous $r_K^{-2}(r)$
in the Limber equation (27). With these factors accounted for, and making the
additional approximation $k \chi \approx k \chi^\prime \ell$ (see
App.~\ref{sec:schmidt08}) we reproduce the expressions of
\cite{2012MNRAS.422.2854G}.
